---
title: "Analysis of Private Healthcare Information Network Data"
author: "Sebastien Peytrignet"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
#Load packages

library(tidyverse)
library(stringr)
library(tidyr)
library(data.table)
library(readxl)
library(writexl)
library(janitor)
library(aws.s3)
library(ggbeeswarm)
library(ggrepel)
library(plotly)
library(ggbreak)

#Clean up the global environment
rm(list = ls())

#Directories in S3
IHT_bucket <- "s3://thf-dap-tier0-projects-iht-067208b7-projectbucket-1mrmynh0q7ljp"
PHIN_subfolder <- "/Private healthcare"
R_workbench <- path.expand("~")
localgit <- dirname(rstudioapi::getSourceEditorContext()$path)

#Import data
source('0. Import data.R')

#Theme
custom_theme <- theme(panel.border = element_blank(),
                      strip.text = element_text(size=10),
                      text = element_text(size = 10),
                      legend.title=element_text(size=10),
                      legend.text=element_text(size=10),
                      axis.text = element_text(size = 10),
                      axis.text.y = element_text(size = 10),
                      axis.text.x = element_text(angle = 45, hjust = 1,size = 10),
                      axis.title.x = element_text(margin = unit(c(3, 0, 0, 0), "mm"),size = 10),
                      axis.title.y = element_text(size = 10))

custom_theme_horiz <- theme(legend.position = "none",
                      panel.border = element_blank(),
                      strip.text = element_text(size=10),
                      text = element_text(size = 10),
                      legend.title=element_text(size=10),
                      legend.text=element_text(size=10),
                      axis.text = element_text(size = 10),
                      axis.text.y = element_text(size = 10),
                      axis.text.x = element_text(angle = 0, hjust = 1,size = 10),
                      axis.title.x = element_text(margin = unit(c(3, 0, 0, 0), "mm"),size = 10),
                      axis.title.y = element_text(size = 10))
```

## Data considerations

- This dataset is provided by the The Private Healthcare Information Network ([PHIN]((https://www.phin.org.uk/data))), who are mandated to report activity data on privately-funded health care under competition and consumer protection laws.
- This dataset also reports activity funded by the NHS for the purpose of comparing NHS-funded to privately-funded activity. Some NHS-funded activity data is in fact delivered by independent sector providers, but this appears as NHS activity in this data because activity is classified by source of funding.
- There have been [concerns](https://www.computerweekly.com/news/252528354/Private-health-provider-data-could-be-shared-with-NHS-England) about private providers submitting [incomplete data](https://www.gov.uk/government/news/action-against-private-hospitals-to-boost-patients-access-to-information) to PHIN. 

<!-- ## Key points -->

<!-- - Therapeutics (chemotherapy, IV drug infusions), upper GI endoscopies, dialysis and colonoscopies are the highest-volume NHS procedures -->
<!-- - The independent sector also does many cataract surgeries, and hip and knee replacements -->
<!-- - In the past year, cataracts, endoscopies, hip replacements and colonoscopies are where the independent sector has added the most treatments -->
<!-- - But this hasn't resulted in large changes in the relative NHS/private mix, because the NHS has also grown in those treatment areas -->
<!-- - In the past two years, specialists (featured in this dataset) are delivering approximately 30% of their treatments in the independent sector -->
<!-- - But there is a lot of variation, with some specialists working almost exclusively in the independent sector -->
<!-- - On average, specialists haven't changed their NHS/private mix over the past two years -->
<!-- - There isn't evidence of certain specialties (e.g. ophthalmologists) 'switching to private' - apart from plastic surgery where the % private went up by 10 points on average in the last year -->
<!-- - The majority of procedures are carried out more quickly in the NHS compared to the independent sector, measured by how many are done as day-cases -->
<!-- - This may be driven by case mix, or because of available resources such as number of beds -->
<!-- - Some procedures have a much shorter length-of-stay in the NHS, e.g. radiotherapy is almost always done as a day-case in NHS but usually with an overnight stay in the independent sector -->
<!-- - On the other hand, some procedures are done a lot more quickly in the independent sector, e.g. weight loss surgery -->
<!-- - For cataracts, endoscopies and joint replacements the NHS/private length-of-stay differential has been stable since 2020/21 -->

## What are the highest-volume NHS and IS/private procedures?

- Therapeutics (chemotherapy, IV drug infusions), upper GI endoscopies, dialysis and colonoscopies are the highest-volume NHS-funded procedures
- The independent sector also does many cataract surgeries, and hip and knee replacements

```{r desc1, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
##############################################
################### By site ##################
##############################################

vol_nation_proc <- plyr::rbind.fill(vol_nation_proc_2122,vol_nation_proc_2021) %>%
  mutate(year=case_when(period_from=="2021-04-01"&period_to=="2022-03-31" ~ "20212022",
                        period_from=="2020-04-01"&period_to=="2021-03-31" ~ "20202021",
                        TRUE ~ "NA"))

vol_nation_proc_clean <- vol_nation_proc %>%
  filter(.,nation=="England") %>% 
  select(.,year,nation,procedure_group,procedure_group_id,private_procedures,nhs_procedures,all_activity_procedures,period_from,period_to) %>%
  mutate_at(c("private_procedures","nhs_procedures"), as.numeric) %>%
  mutate(new_total = rowSums(across(c(private_procedures,nhs_procedures)), na.rm = TRUE)) %>%
  mutate(pct_NHS=nhs_procedures/new_total*100) %>% 
  arrange(desc(new_total))

#Chart 0
#What are private hospitals (incl private wings in NHS) doing? How much of that is paid by NHS?

top10_private_2122 <- vol_nation_proc_clean %>%  filter(year=="20212022") %>% slice_max(private_procedures,n=10) %>% pull(procedure_group)
top10_nhs_2122 <- vol_nation_proc_clean %>%  filter(year=="20212022") %>% slice_max(nhs_procedures,n=10) %>% pull(procedure_group)

cross_bubble_data <- vol_nation_proc_clean %>%
  filter(year=="20212022"&(procedure_group %in% c(top10_nhs_2122,top10_private_2122))) %>%
  select(procedure_group,private_procedures,nhs_procedures) %>%
  pivot_longer(!procedure_group, names_to = "type", values_to = "treatments") %>%
  mutate(list=case_when((procedure_group %in% top10_nhs_2122)&(procedure_group %in% top10_private_2122) ~ "both",
                        (procedure_group %in% top10_nhs_2122)&!(procedure_group %in% top10_private_2122) ~ "nhs",
                        !(procedure_group %in% top10_nhs_2122)&(procedure_group %in% top10_private_2122) ~ "private",
                        TRUE ~ "NA"), 
         type=str_replace_all(type,"_procedures",""),
         nhs=ifelse(type=="nhs",treatments,NA),
         private=ifelse(type=="private",treatments,NA)) %>%
  group_by(procedure_group) %>%
  mutate(total=sum(treatments,na.rm=TRUE),
         nhs=max(nhs,na.rm=TRUE),
         private=max(private,na.rm=TRUE)) %>% 
  ungroup() %>%
  mutate(pct=treatments/total*100) %>% 
  arrange(desc(nhs))

bubble_table <- vol_nation_proc_clean %>%
  filter(year=="20212022"&(procedure_group %in% c(top10_nhs_2122,top10_private_2122))) %>%
  select(procedure_group,private_procedures,nhs_procedures)

#Data table
knitr::kable(bubble_table, align = "lccrr",format.args = list(big.mark = ","))
```

```{r desc2, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=7,fig.height=10}
cross_bubble_nhs <-  filter(cross_bubble_data,list=="nhs"|list=="both") %>%
  ggplot(., aes(x = type,y = reorder(procedure_group,nhs),
                colour = type, size = treatments)) +
  geom_point() +
  geom_text(aes(label = paste0(round(pct,0),"%")), 
            colour = "black", 
            size = 3) +
  scale_x_discrete(position = "top") +
  scale_size_continuous(range = c(5, 25)) + # Adjust as required.
  scale_color_brewer(palette = "Set1") +
  labs(title='Top 10 NHS-funded procedures by volume', subtitle='2021/22') +
  labs(x = NULL, y = NULL) +
  theme_bw() +
  custom_theme_horiz
cross_bubble_nhs
```

```{r desc3, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=7,fig.height=10}
cross_bubble_private <-  filter(cross_bubble_data,list=="private"|list=="both") %>%
  ggplot(., aes(x = type,y = reorder(procedure_group,private),
                colour = type, size = treatments)) +
  geom_point() +
  geom_text(aes(label = paste0(round(pct,0),"%")), 
            colour = "black", 
            size = 3) +
  scale_x_discrete(position = "top") +
  scale_size_continuous(range = c(5, 25)) + # Adjust as required.
  scale_color_brewer(palette = "Set1") +
  labs(title='Top 10 privately-funded procedures by volume', subtitle='2021/22') +
  labs(x = NULL, y = NULL) +
  theme_bw() +
  custom_theme_horiz
cross_bubble_private
```

## Where did the independent sector add the most treatments in 2021/22 compared to 2020/21? Did it disrupt the NHS/private mix for those treatment groups?

- In the past year, cataracts, endoscopies, hip replacements and colonoscopies is where the independent sector has added the most treatments
- But this hasn't resulted in large changes in the relative NHS/private funding mix, because the NHS has also grown in those treatment areas

```{r desc4, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
delta_rankings <- vol_nation_proc_clean %>%
  select(.,year,procedure_group,
         private_procedures,nhs_procedures,new_total) %>%
  mutate(pct_nhs=nhs_procedures/new_total*100) %>%
  pivot_wider(
    names_from = year,
    names_sep = "_",
    values_from = c(new_total,private_procedures,nhs_procedures,pct_nhs)
  ) %>%
  mutate(delta_pct_NHS=pct_nhs_20212022-pct_nhs_20202021,
         delta_all=new_total_20212022-new_total_20202021,
         delta_NHS=nhs_procedures_20212022-nhs_procedures_20202021,
         delta_private=private_procedures_20212022-private_procedures_20202021) %>%
  select(procedure_group,delta_all,delta_NHS,delta_private) %>%
  pivot_longer(!procedure_group, names_to = "type", values_to = "delta") %>%
  filter(type!="delta_all"&!is.na(delta)) %>%
  mutate(abs_delta=abs(delta)) %>% 
  arrange(desc(abs_delta))

table_delta_private <- delta_rankings %>%
  filter(type=="delta_private") %>% 
  slice_max(abs_delta,n=5) %>%
  select(-"abs_delta")

top_private_changes <- delta_rankings %>%
  filter(type=="delta_private") %>% 
  slice_max(abs_delta,n=5) %>%
  pull(procedure_group)

#Data table
knitr::kable(table_delta_private, align = "lccrr")
```

```{r desc5, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10,fig.height=5}
chart1 <- vol_nation_proc_clean %>%
  select(.,year,procedure_group,private_procedures,nhs_procedures) %>%
  filter(procedure_group %in% top_private_changes) %>%
  arrange(procedure_group,year) %>%
  mutate(year=ifelse(year=="20202021","2020/21","2021/22")) %>%
  pivot_longer(!c(year,procedure_group), names_to = "type", values_to = "count") %>%
  group_by(year,procedure_group) %>%
  mutate(new_total=sum(count)) %>% 
  ungroup() %>% 
  mutate(count_pct=ifelse(type=="private_procedures",count/new_total,NA)) %>%
  ggplot(.) +
  facet_wrap(~procedure_group, ncol=3) +
  geom_bar(aes(fill=type, y=count, x=year), position="stack", stat="identity") +
  geom_line(aes(x = year, y = count_pct*1000000, group=type), stat="identity",
            size = 1, col="black") +
  scale_y_continuous(labels = scales::comma, name="Treatments",
                     sec.axis = sec_axis(~./1000000, name = "Share of private procedures",labels = scales::percent)) +
  scale_fill_brewer(palette="Set1",direction=-1) +
  theme_bw() +
  custom_theme
chart1
```

<!-- ## Are consultants doing more private work relative to NHS in 2021/22 compared to 2020/21? -->

<!-- ```{r desc6, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE} -->
<!-- #################################################### -->
<!-- ################### By consultant ################## -->
<!-- #################################################### -->

<!-- #Consultant reference data -->
<!-- #unique(consultants$gmc_specialty_1) %>%  sort() -->

<!-- consultants <- consultants %>% -->
<!--   select(consultant_gmc,starts_with("gmc_specialty")) %>% -->
<!--   splitstackshape::merged.stack(., var.stubs ="gmc_specialty", sep = "_") %>% -->
<!--   select(-c(".time_1")) %>% -->
<!--   dplyr::rename(order='.time_2') %>% -->
<!--   filter(!is.na(gmc_specialty)) %>% -->
<!--   mutate(ophthalmology=ifelse(gmc_specialty %in% c("Ophthalmology","Ophthalmic Surgery"),1,0), -->
<!--          gastroenterology=ifelse(gmc_specialty %in% c("Gastroenterology","Gastro-enterology"),1,0), -->
<!--          orthopaedics=ifelse(gmc_specialty %in% c("Orthopaedics", -->
<!--                                                   "Orthopaedic surgery", -->
<!--                                                   "Trauma and Orthopaedics", -->
<!--                                                   "Trauma and orthopaedic surgery"),1,0), -->
<!--          plastics=ifelse(gmc_specialty %in% c("Plastic surgery"),1,0), -->
<!--          cardiology=ifelse(gmc_specialty %in% c("Cardiology","Cardio-thoracic surgery","Cardiothoracic surgery"),1,0)) %>%  -->
<!--   group_by(consultant_gmc) %>% -->
<!--   summarise(number_specialty=max(order), -->
<!--             first_specialty=first(gmc_specialty), -->
<!--             ophthalmology=max(ophthalmology), -->
<!--             gastroenterology=max(gastroenterology), -->
<!--             orthopaedics=max(orthopaedics), -->
<!--             plastics=max(plastics), -->
<!--             cardiology=max(cardiology)) %>%  -->
<!--   ungroup() %>% -->
<!--   mutate(specialty=ifelse(number_specialty>1,"Multiple",first_specialty)) %>% -->
<!--   select(consultant_gmc,specialty,ophthalmology,gastroenterology,orthopaedics,plastics,cardiology) -->

<!-- #Vol by Nation by Proc -->
<!-- #Cleaning -->

<!-- vol_consultant_proc <- plyr::rbind.fill(vol_consultant_proc_2021,vol_consultant_proc_2122) %>% -->
<!--   left_join(.,consultants,by="consultant_gmc") %>% -->
<!--   mutate(specialty=ifelse(is.na(specialty),"Unknown",specialty)) %>% -->
<!--   mutate(.,private_episodes_num=ifelse(private_episodes=="*","5",private_episodes) %>% as.numeric(), -->
<!--          nhs_episodes_num=ifelse(nhs_episodes=="*","5",nhs_episodes) %>% as.numeric(), -->
<!--          year=case_when(period_from=="2021-04-01"&period_to=="2022-03-31" ~ "20212022", -->
<!--                         period_from=="2020-04-01"&period_to=="2021-03-31" ~ "20202021", -->
<!--                         TRUE ~ "NA")) %>% -->
<!--   mutate(.,total_episodes_num=rowSums(across(ends_with("_num")),na.rm=TRUE), -->
<!--          nhs_to_private_episodes=nhs_episodes_num/private_episodes_num) %>% -->
<!--   mutate( -->
<!--     pct_private=private_episodes_num/total_episodes_num*100, -->
<!--     more_ratio=ifelse(nhs_to_private_episodes<1,"more private","more NHS")) -->

<!-- # vol_consultant_proc %>% -->
<!-- #   filter(nhs_episodes_num==0|private_episodes_num==0) -->

<!-- vol_consultant_proc_wide <- vol_consultant_proc %>% -->
<!--   select(.,consultant_gmc,consultant_forenames,consultant_surname,specialty, -->
<!--          ophthalmology,gastroenterology,orthopaedics,plastics,cardiology, -->
<!--          pct_private,nhs_to_private_episodes,nhs_episodes_num,private_episodes_num, -->
<!--          total_episodes_num,year) %>% -->
<!--   filter(!is.na(nhs_to_private_episodes)) %>% -->
<!--   pivot_wider(names_from = year, -->
<!--               names_sep = "_", -->
<!--               values_from = c(pct_private,nhs_to_private_episodes, private_episodes_num, nhs_episodes_num,total_episodes_num)) %>% -->
<!--   mutate(pct_private_diff=pct_private_20212022-pct_private_20202021, -->
<!--          ratio_diff=nhs_to_private_episodes_20212022-nhs_to_private_episodes_20202021, -->
<!--          ratio_direction=ifelse(nhs_to_private_episodes_20212022>nhs_to_private_episodes_20202021, -->
<!--                                 "more NHS","more private")) %>% -->
<!--   mutate(specsum=rowSums(across(c("ophthalmology","gastroenterology","orthopaedics","plastics","cardiology")))) %>% -->
<!--   mutate(.,spec_catgory_bis=case_when(specsum>1 ~ "multiple main specialties", -->
<!--                                       ophthalmology==1 ~ "ophthalmology", -->
<!--                                       gastroenterology==1 ~ "gastroenterology", -->
<!--                                       orthopaedics==1 ~ "orthopaedics", -->
<!--                                       plastics==1 ~ "plastics", -->
<!--                                       cardiology==1 ~ "cardiology", -->
<!--                                       TRUE ~ "other")) -->
<!-- ``` -->

<!-- ```{r desc7, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10,fig.height=5} -->
<!-- chart3_data <- vol_consultant_proc %>% -->
<!--   mutate(year=ifelse(year=="20202021","2020/21","2021/22")) %>% -->
<!--   group_by(year) %>% -->
<!--   summarise(pct_private=mean(pct_private,na.rm=TRUE)) %>%  -->
<!--   ungroup() -->

<!-- #Data table -->
<!-- knitr::kable(chart3_data, align = "lccrr") -->

<!-- #Chart -->
<!-- chart3 <- vol_consultant_proc %>% -->
<!--   mutate(year=ifelse(year=="20202021","2020/21","2021/22")) %>%  -->
<!--   ggplot(., aes(x=pct_private)) + -->
<!--   facet_wrap(~year, scales = "free") + -->
<!--   geom_histogram(fill="aquamarine", position="dodge",col="black")+ -->
<!--   xlab("Percent of treatments delivered privately") + -->
<!--   ylab("Number of consultants") + -->
<!--   theme_bw() + -->
<!--   custom_theme -->
<!-- chart3 -->
<!-- ``` -->

<!-- ```{r desc8, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10,fig.height=5} -->
<!-- chart5 <- vol_consultant_proc_wide %>%  -->
<!--   ggplot(., aes(x=pct_private_diff)) + -->
<!--   geom_histogram(fill="tomato", position="dodge",col="black")+ -->
<!--   geom_vline(xintercept =0,col="red",lwd=1.5) + -->
<!--   xlab("Change in percentage private (points)") + -->
<!--   ylab("Number of consultants") + -->
<!--   annotate("text", x = 40, y = 200, size = 3, label = "Did more private →") + -->
<!--   annotate("text", x = -40, y = 200, size = 3, label = "← Did more NHS") + -->
<!--   theme_bw() + -->
<!--   custom_theme -->
<!-- chart5 -->
<!-- ``` -->

<!-- ```{r desc9, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10,fig.height=5} -->
<!-- mean_pct_diff_by_main_specialty_bis <- vol_consultant_proc_wide %>% -->
<!--   group_by(spec_catgory_bis) %>% -->
<!--   summarise(pct_private_diff=mean(pct_private_diff,na.rm=TRUE), -->
<!--             consultants=n(), -->
<!--             total_episodes_num_20212022=sum(total_episodes_num_20212022,na.rm=TRUE)) %>%  -->
<!--   ungroup() %>% -->
<!--   arrange(desc(pct_private_diff)) -->

<!-- #Data table -->
<!-- knitr::kable(mean_pct_diff_by_main_specialty_bis, align = "lccrr") -->

<!-- #Chart -->
<!-- beeswarm <- ggplot(vol_consultant_proc_wide, aes(x = spec_catgory_bis, y = pct_private_diff, fill = spec_catgory_bis)) + -->
<!--   geom_violin(alpha = 0.5) + -->
<!--   xlab("Main specialty") + -->
<!--   ylab("Change in percentage private (points)") + -->
<!--   annotate("text", x = 1, y = 45, size = 3, label = "Did more private ↑") + -->
<!--   annotate("text", x = 1, y = -40, size = 3, label = "Did more NHS ↓") + -->
<!--   theme_bw() + -->
<!--   custom_theme + -->
<!--   theme(legend.position = "none") -->
<!-- beeswarm -->
<!-- ``` -->

## Complexity: do independent sector patients have shorter stays compated to NHS funded? 

- The majority of procedures are carried out more quickly in the NHS compared to the independent sector, measured by how many are done as day-cases
- This may be driven by case mix, or because of available resources or number of beds

```{r desc10, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
los_nation_proc <- plyr::rbind.fill(los_nation_proc_2122,los_nation_proc_2021) %>%
  mutate(year=case_when(period_from=="2021-04-01"&period_to=="2022-03-31" ~ "20212022",
                        period_from=="2020-04-01"&period_to=="2021-03-31" ~ "20202021",
                        TRUE ~ "NA"))

los_nation_proc <- los_nation_proc %>%
  filter(.,nation=="England") %>% 
  select(.,year,nation,procedure_group,procedure_group_id,starts_with("private"),starts_with("nhs"),starts_with("all_activity"),period_from,period_to) %>%
  mutate_at(c("private_inpatient_median_los","nhs_inpatient_median_los",
              "private_daycase_rate_percent","nhs_daycase_rate_percent",
              "private_inpatient_spells","nhs_inpatient_spells","all_activity_inpatient_spells"), as.numeric) %>%
  mutate(pct_private=private_inpatient_spells/all_activity_inpatient_spells*100)

los_nation_proc_wide <- los_nation_proc %>%
  mutate(spread_dayrate=private_daycase_rate_percent-nhs_daycase_rate_percent,
         spread_median=private_inpatient_median_los-nhs_inpatient_median_los) %>%
  mutate(spread_direction=ifelse(spread_dayrate>0,"more day-cases in private","more day-cases in NHS")) %>% 
  select(year,procedure_group,private_inpatient_spells,nhs_inpatient_spells,all_activity_inpatient_spells,
         private_daycase_rate_percent,nhs_daycase_rate_percent,spread_dayrate,spread_median,
         pct_private,spread_direction) %>%
  pivot_wider(names_from = year,
              names_sep = "_",
              values_from = c(private_inpatient_spells,nhs_inpatient_spells,all_activity_inpatient_spells,
                              private_daycase_rate_percent,nhs_daycase_rate_percent,
                              spread_dayrate,spread_median,pct_private,spread_direction)) %>%
  mutate(spread_dayrate_delta=spread_dayrate_20212022-spread_dayrate_20202021,
         spread_median_delta=spread_median_20212022-spread_median_20202021)

static_spread_ranking <- los_nation_proc_wide %>%
  filter(!is.na(spread_dayrate_20212022)) %>%
  select(procedure_group,spread_dayrate_20212022,private_daycase_rate_percent_20212022,
         nhs_daycase_rate_percent_20212022,private_inpatient_spells_20212022,
         nhs_inpatient_spells_20212022,all_activity_inpatient_spells_20212022,
         pct_private_20212022,spread_direction_20212022) %>% 
  mutate(abs_spread_dayrate_20212022=abs(spread_dayrate_20212022)) %>% 
  arrange(desc(abs_spread_dayrate_20212022))
```

```{r desc11, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10,fig.height=5}
barchart_comparison <- static_spread_ranking %>%
  tabyl(spread_direction_20212022) %>% 
  ggplot(data=., aes(x=spread_direction_20212022, y=n, fill=spread_direction_20212022)) +
  geom_bar(stat="identity")+
  geom_text(aes(label=n), vjust=1.6, color="white", size=3.5)+
  theme_bw() +
  labs(title='Day-cases in NHS vs private', subtitle='2021/22, by procedures') +
  xlab(" ") +
  ylab("Procedures") +
  theme(legend.position = "none") +
  scale_fill_brewer(palette="Set1",direction=-1) +
  theme_bw() +
  theme(legend.position = "none")
barchart_comparison
```

## Which procedures does the NHS do quicker?

- Some procedures have a much shorter length-of-stay in the NHS, e.g. radiotherapy is almost always done as a day-case in NHS but usually with an overnight stay in the independent sector

```{r desc12, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10,fig.height=5}
#Bottom
barbell_data_bottom <- static_spread_ranking %>%
  slice_min(spread_dayrate_20212022,n=10) %>%
  arrange(desc(spread_dayrate_20212022)) %>% 
  select(procedure_group,spread_dayrate_20212022,nhs_daycase_rate_percent_20212022,private_daycase_rate_percent_20212022) %>%
  pivot_longer(!c(procedure_group,spread_dayrate_20212022), names_to = "sector", values_to = "pct_day_case") %>%
  mutate(sector_name=case_when(sector=="nhs_daycase_rate_percent_20212022" ~ "NHS",
                               sector=="private_daycase_rate_percent_20212022" ~ "Private",
                               TRUE ~ "NA")) %>%
  select(-c("sector"))
barbell_data_bottom_nhs <- barbell_data_bottom %>%  filter(sector_name=="NHS")
barbell_data_bottom_private <- barbell_data_bottom %>%  filter(sector_name=="Private")

barbell_bottom <-  ggplot(barbell_data_bottom) +
  geom_segment(data=barbell_data_bottom_nhs,
               aes(x = pct_day_case, y = reorder(procedure_group,-pct_day_case),
                   yend = barbell_data_bottom_private$procedure_group, xend = barbell_data_bottom_private$pct_day_case),
               color = "#aeb6bf",
               size = 4.5,
               alpha = .5) +
  geom_point(aes(x = pct_day_case, y = reorder(procedure_group,-pct_day_case), color = sector_name), size = 4, show.legend = TRUE)+
  theme_bw() +
  labs(title='Procedures the NHS does quicker', subtitle='2021/22, top 10 largest gap') +
  xlab("Percentage of day-cases") +
  ylab("Procedure") +
  scale_color_brewer(name="Sector",palette="Set1",direction=-1)
barbell_bottom
```

## Which procedures does the independent sector do quicker?

- Some procedures are done a lot more quickly in the independent sector, e.g. weight loss surgery.

```{r desc13, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10,fig.height=5}
barbell_data_top <- static_spread_ranking %>%
  slice_max(spread_dayrate_20212022,n=10) %>%
  arrange(desc(spread_dayrate_20212022)) %>% 
  select(procedure_group,spread_dayrate_20212022,nhs_daycase_rate_percent_20212022,private_daycase_rate_percent_20212022) %>%
  pivot_longer(!c(procedure_group,spread_dayrate_20212022), names_to = "sector", values_to = "pct_day_case") %>%
  mutate(sector_name=case_when(sector=="nhs_daycase_rate_percent_20212022" ~ "NHS",
                               sector=="private_daycase_rate_percent_20212022" ~ "Private",
                               TRUE ~ "NA")) %>%
  select(-c("sector"))
barbell_data_top_nhs <- barbell_data_top %>%  filter(sector_name=="NHS")
barbell_data_top_private <- barbell_data_top %>%  filter(sector_name=="Private")

barbell_top <-  ggplot(barbell_data_top) +
  geom_segment(data=barbell_data_top_nhs,
               aes(x = pct_day_case, y = reorder(procedure_group,-pct_day_case),
               yend = barbell_data_top_private$procedure_group, xend = barbell_data_top_private$pct_day_case), 
               color = "#aeb6bf",
               size = 4.5,
               alpha = .5) +
  geom_point(aes(x = pct_day_case, y = reorder(procedure_group,-pct_day_case), color = sector_name), size = 4, show.legend = TRUE)+
  theme_bw() +
  labs(title='Procedures the private sector does quicker', subtitle='2021/22, top 10 largest gap') +
  xlab("Percentage of day-cases") +
  ylab("Procedure") +
  scale_color_brewer(name="Sector",palette="Set1",direction=-1)
barbell_top
```

<!-- ## Has the length-of-stay gap between NHS and independent sector widened between 2020/21 and 2021/22? (selection of procedures) -->

<!-- - For cataracts, endoscopies and joint replacements the NHS/private length-of-stay differential has been stable since 2020/21. -->

<!-- ```{r desc14, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10,fig.height=5} -->
<!-- #Selection of longitudinal -->
<!-- barbell_data_selection <- los_nation_proc %>% -->
<!--   mutate(spread_dayrate=private_daycase_rate_percent-nhs_daycase_rate_percent, -->
<!--          year=case_when(year=="20212022" ~ "2021/22", -->
<!--                         year=="20202021" ~ "2020/21", -->
<!--                         TRUE ~ "NA")) %>% -->
<!--   mutate(procedure_year=paste(procedure_group,year,sep=" ")) %>% -->
<!--   filter(procedure_group %in% c("Cataract surgery", -->
<!--                                 "Hip replacement (primary)", -->
<!--                                 "Knee replacement (primary)", -->
<!--                                 "Diagnostic upper GI endoscopy")) %>%  -->
<!--   select(procedure_year,spread_dayrate,private_daycase_rate_percent,nhs_daycase_rate_percent) %>% -->
<!--   pivot_longer(!c(procedure_year,spread_dayrate), names_to = "sector", values_to = "pct_day_case") %>% -->
<!--   mutate(sector_name=case_when(sector=="nhs_daycase_rate_percent" ~ "NHS", -->
<!--                                sector=="private_daycase_rate_percent" ~ "Private", -->
<!--                                TRUE ~ "NA")) %>% -->
<!--   select(-c("sector")) -->
<!-- barbell_data_selection_nhs <- barbell_data_selection %>%  filter(sector_name=="NHS") -->
<!-- barbell_data_selection_private <- barbell_data_selection %>%  filter(sector_name=="Private") -->
<!-- # diff_selection <- barbell_data_selection %>%  -->
<!-- #   filter(sector_name == "NHS") %>% -->
<!-- #   mutate(x_pos = as.numeric(pct_day_case)+as.numeric(spread_dayrate)/2) -->

<!-- barbell_selection <-  ggplot(barbell_data_selection) + -->
<!--   geom_segment(data=barbell_data_selection_nhs, -->
<!--                aes(x = pct_day_case, y = reorder(procedure_year,desc(procedure_year)), -->
<!--                    yend = barbell_data_selection_private$procedure_year, xend = barbell_data_selection_private$pct_day_case), -->
<!--                color = "#aeb6bf", -->
<!--                size = 4.5, -->
<!--                alpha = .5) + -->
<!--   geom_point(aes(x = pct_day_case, y = reorder(procedure_year,desc(procedure_year)), color = sector_name), size = 4, show.legend = TRUE)+ -->
<!--   theme_bw() + scale_x_break(c(5, 95)) + -->
<!--   labs(title='Change in day-rate', subtitle='2020/21 and 2021/22, selected procedures') + -->
<!--   xlab("Percentage of day-cases") + -->
<!--   ylab("Procedure/year") + -->
<!--   scale_color_brewer(name="Sector",palette="Set1",direction=-1) -->
<!-- barbell_selection -->
<!-- ``` -->

<!-- ## Are procedures with a wider length-of-stay gap more likely to delivered by the independent sector? -->

<!-- - There is no evidence that procedures where the independent sector is 'quicker' are more likely to be done privately -->

<!-- ```{r desc15, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10,fig.height=5} -->
<!-- static_spread_ranking_highlight <- static_spread_ranking %>% -->
<!--   filter(procedure_group %in% c("Cataract surgery", -->
<!--                                 "Hip replacement (primary)", -->
<!--                                 "Knee replacement (primary)", -->
<!--                                 "Diagnostic upper GI endoscopy")) -->

<!-- bubble_chart_mini <- static_spread_ranking %>% -->
<!--   filter(pct_private_20212022<50) %>%  -->
<!--   ggplot(., aes(x=spread_dayrate_20212022, y=pct_private_20212022, size = all_activity_inpatient_spells_20212022)) + -->
<!--   geom_point(alpha=0.5, shape=21, color="black") + -->
<!--   geom_point(data=static_spread_ranking_highlight, -->
<!--              aes(x=spread_dayrate_20212022, y=pct_private_20212022,fill=procedure_group),alpha=0.8, shape=21) + -->
<!--   geom_label_repel(aes(label = procedure_group), -->
<!--                    nudge_x = 2, -->
<!--                    nudge_y = 1, -->
<!--                    na.rm = TRUE,size=3,alpha=0.7) + -->
<!--   scale_size(range = c(.1, 12), name="Total spells\n(NHS and private)") + -->
<!--   scale_fill_brewer(name="Procedures",palette="Dark2") + -->
<!--   xlim(c(-85,85)) + -->
<!--   theme_bw() + -->
<!--   labs(title='Complexity of NHS vs. private procedures', subtitle='2021/22') + -->
<!--   xlab("Spread in day-case rate\n(% day-cases private - % day-cases NHS)") + -->
<!--   ylab("Percent private") + -->
<!--   annotate("text", x = 45, y = 15, size = 3, label = "More day-cases in private →") + -->
<!--   annotate("text", x = -45, y = 15, size = 3, label = "← More day-cases in NHS") -->
<!-- bubble_chart_mini -->
<!-- ``` -->